---
title: "OC Stemr Update"
author: "Damon Bayer"
date: "10/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(stemr)
library(coda)
library(tidyverse)
library(cowplot)
source("stemr_functions.R")
source("~/Documents/uci_covid_modeling/code/SEIeIpRD/plot_functions.R")
theme_set(theme_cowplot())
```


```{r build stemr model}
# Read Data ---------------------------------------------------------------
dat <- read_rds("~/Documents/uci_covid_modeling/code/SEIeIpRD/oc/2020-08-12_2020-09-16/model_objects.rds")$lumped_ochca_covid
init_states <- c(S = 3012564L, E = 14866L, Ie = 20907L, Ip = 20590L, R = 0L, D = 0L)
popsize <- sum(init_states)

# Build stem_object -------------------------------------------------------
obs_times <- as.numeric((dat$end_date - min(dat$start_date) + 1L) / 7)

strata <- NULL # no strata
compartments <- c("S", "E", "Ie", "Ip", "R", "D")

rates <-
  list(rate(rate = "beta * (Ie + 0.8 * Ip)", # individual level rate (unlumped)
            from = "S",        # source compartment
            to   = "E",        # destination compartment
            incidence = F),    # compute incidence of S2I transitions, required for simulating incidence data
       rate(rate = "gamma",
            from = "E",
            to = "Ie",
            incidence = F),
       rate(rate = "nu_early",
            from = "Ie",
            to = "Ip",
            incidence = T),
       rate(rate = "mu_rec",
            from = "Ip",
            to = "R",
            incidence = F),
       rate(rate = "mu_death",       # individual level rate
            from = "Ip",        # source compartment
            to   = "D",        # destination compartment
            incidence = TRUE)) # compute incidence of I2R transitions (not required for simulating data)

state_initializer <-
  list(stem_initializer(
    init_states = init_states, # must match compartment names
    fixed = T,
    # prior = init_states / C,
    dist = "dirmultinom"
  )) # initial state fixed for simulation, we'll change this later

parameters <- numeric(10); names(parameters) <- c("beta", "gamma", "nu_early", "mu_rec", "mu_death", "rho_death", "phi_death", "alpha0", "alpha1", "kappa")
constants <- c(t0 = 0)
tcovar <- data.frame(time = obs_times,
                     tests = dat$new_tests)
tmax <- max(tcovar$time)

# list of emission distribution lists (analogous to rate specification)
emissions <-
  list(emission(meas_var = "cases", # transition or compartment being measured (S->I transitions)
                distribution    = "betabinomial",        # emission distribution
                emission_params =
                  c("tests",
                    "kappa * (exp(alpha0) * (Ie2Ip / popsize) ^ alpha1) / (exp(alpha0) * (Ie2Ip / popsize) ^ alpha1 + (1 - Ie2Ip/popsize) ^ alpha1)",
                    "kappa * ((1 - Ie2Ip/popsize) ^ alpha1) / (exp(alpha0) * (Ie2Ip / popsize) ^ alpha1 + (1 - Ie2Ip/popsize) ^ alpha1)"), # distribution pars, here overdispersion and mean
                incidence       = TRUE,                  # is the data incidence
                obstimes        = obs_times), # vector of observation times
       emission(meas_var = "deaths",
                distribution = "negbinomial",
                emission_params = c("phi_death",
                                    "rho_death * Ip2D"),
                incidence = T,
                obstimes        = obs_times)) # vector of observation times)
# list of emission distribution lists (analogous to rate specification)

dynamics <-
  stem_dynamics(
    rates = rates,
    parameters = parameters,
    state_initializer = state_initializer,
    compartments = compartments,
    constants = constants,
    tcovar = tcovar,
    tmax = tmax,
    compile_ode = T,   # compile ODE functions
    compile_rates = F, # compile MJP functions for Gillespie simulation
    compile_lna = T,   # compile LNA functions
    messages = F       # don't print messages
  )



measurement_process <-
  stem_measure(emissions = emissions,
               dynamics = dynamics,
               data = dat %>%
                 mutate(t= obs_times) %>%
                 select(t, cases = new_cases, deaths = new_deaths))

stem_object <- make_stem(dynamics = dynamics, measurement_process = measurement_process)
```




# Overview

```{r}
model_objects <- read_rds("fixed_init_sim/model_objects.rds")

oc_post_stemr_params <- map(fixed_init_multi_chain_stem_fit$stem_fit_list,
function(stem_fit) {
  stem_fit$results$posterior$parameter_samples_nat %>%
    split(., row(.)) %>%
    map(~`names<-`(.,c("beta", "gamma", "nu_early", "mu_rec", "mu_death", "rho_death",
                       "phi_death", "alpha0", "alpha1", "kappa")))
}) %>%
  unlist(recursive = F)
```

```{r}
stemr_prior <- read_rds("priors_only_sim/priors_only_multi_chain_stem_fit.rds") %>% 
  extract_stem_parameter_posterior() %>%
  tidy_draws() %>%
  select(-starts_with(".")) %>%
  mutate(model = "stemr", type = "prior")

stemr_post <- read_rds("fixed_init_sim/fixed_init_multi_chain_stem_fit.rds") %>% 
  extract_stem_parameter_posterior() %>%
  tidy_draws() %>%
  select(-starts_with(".")) %>%
  mutate(model = "stemr", type = "post")

stan_post <- read_rds("fixed_init_sim/oc_post.rds") %>%
  spread_draws(alpha_incid_0,
               alpha_incid_1,
               early_dur,
               IFR,
               latent_dur,
               prog_dur,
               prog_dur,
               sqrt_kappa_inv_incid,
               sqrt_phi_inv_death,
               log_R0,
               rho_death) %>%
  transmute(R0_est = log_R0,
            dur_latent_est = -log(latent_dur),
            dur_early_est = -log(early_dur),
            dur_progress_est = -log(prog_dur),
            ifr_est = logit(IFR),
            rho_death_est = logit(rho_death),
            phi_death_est = log(sqrt_phi_inv_death),
            alpha0_est = log(alpha_incid_0),
            alpha1_est = logit(alpha_incid_1),
            kappa_est = log(sqrt_kappa_inv_incid)) %>%
  mutate(model = "stan", type = "post")

stan_prior <- read_rds("fixed_init_sim/oc_prior.rds") %>%
  spread_draws(alpha_incid_0,
               alpha_incid_1,
               early_dur,
               IFR,
               latent_dur,
               prog_dur,
               prog_dur,
               sqrt_kappa_inv_incid,
               sqrt_phi_inv_death,
               log_R0,
               rho_death) %>%
  transmute(R0_est = log_R0,
            dur_latent_est = log(latent_dur),
            dur_early_est = log(early_dur),
            dur_progress_est = log(prog_dur),
            ifr_est = logit(IFR),
            rho_death_est = logit(rho_death),
            phi_death_est = log(sqrt_phi_inv_death),
            alpha0_est = log(alpha_incid_0),
            alpha1_est = logit(alpha_incid_1),
            kappa_est = log(sqrt_kappa_inv_incid)) %>%
  mutate(model = "stan", type = "prior")
```

```{r}
bind_rows(stemr_prior, stemr_post, stan_prior, stan_post) %>%
  unite(type, model, col = "source") %>%
  pivot_longer(-source) %>%
  group_by(source, name) %>%
  median_qi(.width = c(0.5, 0.95)) %>%
  ggplot(aes(x = value, xmin = .lower, xmax = .upper, color = source)) +
  geom_pointinterval(position = position_dodge(width = .3)) +
  facet_wrap(. ~ name, scales = "free_x") +
  cowplot::theme_minimal_vgrid() +
  xlab(NULL) +
  scale_y_continuous(labels = NULL) +
  ggtitle("Stan / Stemr Comparison with Fixed Initial Compartments") +
  theme(legend.position = c(.6, .2)) +
  guides(colour = guide_legend(reverse = T))
```

Stan and Stemr Priors agree, while the Posteriors disagree to varying levels. Kappa and R0 are especially concerning.

SHOULD ALSO SHOW EPI CURVES

# ODE Solver Check

```{r}
combined_epi_curves <- read_rds("ode_solver_check_sim/combined_epi_curves.rds")
combined_epi_curves %>%
  ggplot(aes(time, value, group = model, color = model)) +
  geom_line(size = 1.25, alpha = 0.5) +
  facet_wrap(. ~ name, scales = "free_y")
```

Here, I give stemr and stan the same parameter values and solve the ODEs. Results are indistinguishable.

# Prior Predictives
```{r}
plot_death_stemr_stan <- function(stemr_rds, stan_rds, title = "Preds") {
  stemr_prior_pred <- map(read_rds(stemr_rds)$stem_fit_list,
                        function(stem_fit) {
                          stem_fit$results$posterior$parameter_samples_nat %>%
                            split(., row(.)) %>%
                            map(~`names<-`(.,c("beta", "gamma", "nu_early", "mu_rec", "mu_death", "rho_death",
                       "phi_death", "alpha0", "alpha1", "kappa")))
                          }) %>%
  unlist(recursive = F) %>% 
  simulate_stem(stem_object = stem_object, method = "ode", nsim = 8000,
                simulation_parameters = .) %>%
  `$`(dataset) %>%
  map_dfr(as_tibble) %>% 
  mutate(t = model_objects$dates_pp[time * 7 / 3]) %>% 
  select(deaths, cases, t)

prior_pred_death <- prep_death(model_objects = model_objects, oc_post = read_rds(stan_rds))

prior_pred_death$stemr <- stemr_prior_pred %>% select(stemr = deaths, t) %>% group_by(t) %>% mutate(.draw = row_number())
prior_pred_death$stan <- prior_pred_death$posterior %>% rename(stan = deaths) %>% filter(t %in% prior_pred_death$stemr$t) %>% group_by(t) %>% mutate(.draw = row_number())
prior_pred_death$posterior <- NULL

my_plot <- ggplot() + 
geom_lineribbon(data = left_join(prior_pred_death$stan, prior_pred_death$stemr) %>% 
  select(-.draw) %>% 
  pivot_longer(-t) %>% 
  group_by(t, name) %>% 
  median_qi(.width = c(0.5, 0.8, 0.95)),
  mapping = aes(t, value, ymin = .lower, ymax = .upper)) +
  facet_wrap(. ~ name) +
  geom_lineribbon() +
  scale_fill_brewer() +
  geom_point(data = prior_pred_death$observed, mapping = aes(t, observed)) +
  ggtitle(title) +
  ylab("Deaths")
  theme(legend.position = "none")
  
  my_plot
}

plot_pos_stemr_stan <- function(stemr_rds, stan_rds, title = "Preds") {
  stemr_prior_pred <- map(read_rds(stemr_rds)$stem_fit_list,
                          function(stem_fit) {
                            stem_fit$results$posterior$parameter_samples_nat %>%
                              split(., row(.)) %>%
                              map(~`names<-`(.,c("beta", "gamma", "nu_early", "mu_rec", "mu_death", "rho_death",
                                                 "phi_death", "alpha0", "alpha1", "kappa")))
                            }) %>%
    unlist(recursive = F) %>% 
    simulate_stem(stem_object = stem_object, method = "ode", nsim = 8000,
                  simulation_parameters = .) %>%
    `$`(dataset) %>%
    map_dfr(as_tibble) %>% 
    mutate(t = model_objects$dates_pp[time * 7 / 3]) %>% 
    select(deaths, cases, t)

  prior_pred_pos <- prep_pos(model_objects = model_objects, oc_post = read_rds(stan_rds))

  prior_pred_pos$stemr <- stemr_prior_pred %>% 
    select(t, cases) %>% 
    left_join(select(model_objects$lumped_ochca_covid, t = end_date, tests = new_tests)) %>% 
    mutate(stemr = cases / tests) %>% 
    select(t, stemr) %>% 
    group_by(t) %>%
    mutate(.draw = row_number())
  
  prior_pred_pos$stan <- prior_pred_pos$posterior %>% rename(stan = percent_positive) %>% filter(t %in% prior_pred_pos$stemr$t) %>% group_by(t) %>% mutate(.draw = row_number())
  prior_pred_pos$posterior <- NULL

    my_plot <- ggplot() + 
      geom_lineribbon(data = left_join(prior_pred_pos$stan, prior_pred_pos$stemr) %>% 
      select(-.draw) %>% 
      pivot_longer(-t) %>% 
      group_by(t, name) %>% 
      median_qi(.width = c(0.5, 0.8, 0.95)),
      mapping = aes(t, value, ymin = .lower, ymax = .upper)) +
      facet_wrap(. ~ name) +
      geom_lineribbon() +
      scale_fill_brewer() +
      geom_point(data = prior_pred_pos$observed, mapping = aes(t, observed)) +
      ggtitle(title) +
      ylab("Positivity")
      theme(legend.position = "none")
  
  my_plot
}
```


```{r}
prior_death_pred_plot <- plot_death_stemr_stan(stemr_rds = "priors_only_sim/priors_only_multi_chain_stem_fit.rds", stan_rds = "fixed_init_sim/oc_prior.rds", "Prior Predictive")
prior_death_pred_plot

prior_pos_pred_plot <- plot_pos_stemr_stan(stemr_rds = "priors_only_sim/priors_only_multi_chain_stem_fit.rds", stan_rds = "fixed_init_sim/oc_prior.rds", "Prior Predictive")
prior_pos_pred_plot
```


# Posterior Predictives

```{r}
post_death_pred_plot <- plot_death_stemr_stan(stemr_rds = "fixed_init_sim/fixed_init_multi_chain_stem_fit.rds", stan_rds = "fixed_init_sim/oc_post.rds", "Posterior Predictive")
post_death_pred_plot

post_pos_pred_plot <- plot_pos_stemr_stan(stemr_rds = "fixed_init_sim/fixed_init_multi_chain_stem_fit.rds", stan_rds = "fixed_init_sim/oc_post.rds", "Posterior Predictive")
post_pos_pred_plot
```
From the previous plots, stemr has higher log(1 / sqrt(kappa)) so "more overdispersion" which is exactly what we see here.

# Stemr Sampler
